---
title: "adjusted MIC"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# R-code to estimate the adjusted MIC, predictive MIC and ROC-based MIC  

This code simulates an example data set with 10 items that are measured at two time points. 

```{r}
# Acquire packages
library(mirt)     # for simulating data
library(lavaan)   # for estimating the transition ratings reliability
library(pROC)     # for estimating ROC-based MIC
``` 

## Simulate dataset using 'mirt'

### Gernerate a set of IRT item parameters

```{r}

set.seed(12345)

b2  <- c(-0.8, -0.8, -0.4, -0.4, 0, 0, 0.4, 0.4, 0.8, 0.8)
bc <- b2/4
b1 <- b2 - 1 + sample(bc)
b3 <- b2 + 1 + sample(bc)
a1 <- sample( 1+b2/2 )

( cf.simb <- data.frame(a1,b1,b2,b3) )
``` 


### Transform b-parameters to d-parameters (mirt works with d-parameters)

difficulty (b) = easiness (d) / -a

```{r}
cf.sim <- cf.simb

colnames(cf.sim) <- c("a1","d1","d2","d3")

cf.sim$d1 <- -cf.simb$b1*cf.sim$a1
cf.sim$d2 <- -cf.simb$b2*cf.sim$a1
cf.sim$d3 <- -cf.simb$b3*cf.sim$a1
round(cf.sim, 3)
round(colMeans(cf.sim), 3)
```



### Create theta T1 and dataset 1 (baseline)

```{r}
a1 <- as.matrix(cf.sim[ , 1])
d1 <- as.matrix(cf.sim[ , -1])

N <- 1000         # sample size
prop.imp <- 0.3   # proportion improved
mn.imic  <- 0.5   # mean individual MICs in terms of theta change

tet1s <- rnorm(N, 0, 1)   # simulate theta T1
tet1s <- as.matrix(tet1s)

set.seed( sample(10000:20000,1) )

dat1 <- simdata(a1, d1, N, itemtype="graded", Theta=tet1s)
dat1 <- as.data.frame(dat1)

xo1 <- rowSums(dat1)   # This is the baseline test (sum) score
```

### Create theta change

```{r}
rt1ch <- -0.5            # correlation between theta T1 and theta change
tetchs <- rt1ch*tet1s + sqrt(1-rt1ch^2)*rnorm(N, mean(tet1s), sd(tet1s))
cor(tet1s,tetchs)

tetchs <- tetchs - mean(tetchs)         # makes mean = 0
tetchs <- (tetchs/sd(tetchs))*1         # makes SD = 1

( qtl <- quantile(tetchs, prob=(1-prop.imp)) )
( mean.tetchs <- mn.imic - qtl )        # Estimate mean theta change to get
# the desired proportion improved
tetchs <- tetchs + mean.tetchs          # transform mean of tetchs
```

### Create theta T2 and dataset 2  (follow-up)

```{r}
tet2s <- as.matrix(tet1s + tetchs)     # theta T2

cor(tet1s, tet2s)   

set.seed( sample(30000:40000,1) )

dat2 <- simdata(a1, d1, N, itemtype="graded", Theta=tet2s)
dat2 <- as.data.frame(dat2)
```

### Insert error in transition ratings

```{r}

rel.trat <- 0.5   # reliability of the perceived change/transition rating

( sd.ch.error <- sqrt(((1-rel.trat)/rel.trat)*sd(tetchs)^2) )
tetch.percv <- tetchs + rnorm(N, 0, sd.ch.error)  # perceived change
```

### Create iMIC distribution
```{r}
imic <- rnorm(N, mn.imic, 0.1*mn.imic)
```

### create dichotomous transition ratings

```{r}
trat <- numeric(N)                         
trat[tetch.percv > imic] <- 1   
mean(trat)                           # proportion improved based on perceived change
```

### Create dataset

```{r}
org <- data.frame(dat1, dat2, trat)
head(org)
nitems = 10     # Provide the number of items in the scale

## Simplify/standardize the item names
names(org)[1:nitems] <- paste0('v1', '_', 1:nitems)
names(org)[(nitems+1):(2*nitems)] <- paste0('v2', '_', 1:nitems)
names(org)[2*nitems+1] <- "trat"
names(org)

## copy original data into a workfile (dat)
dat <- org
head(dat)

```




## Estimate TR reliability with lavaan

### Specify the factor model

A factor for each time point, with correlated errors over time.

```{r}
model <- ' 
   # Factors
   F1 =~ a1*v1_1+a2*v1_2+a3*v1_3+a4*v1_4+a5*v1_5+a6*v1_6+a7*v1_7+a8*v1_8+
         a9*v1_9+a10*v1_10+trat
   F2 =~ a1*v2_1+a2*v2_2+a3*v2_3+a4*v2_4+a5*v2_5+a6*v2_6+a7*v2_7+a8*v2_8+
         a9*v2_9+a10*v2_10+trat
   
   # Correlated errors over time
   v1_1 ~~ v2_1
   v1_2 ~~ v2_2
   v1_3 ~~ v2_3
   v1_4 ~~ v2_4
   v1_5 ~~ v2_5
   v1_6 ~~ v2_6
   v1_7 ~~ v2_7
   v1_8 ~~ v2_8
   v1_9 ~~ v2_9
   v1_10 ~~ v2_10
   '
```

### Fit the confirmatory factor analysis

```{r}
fit <- cfa(model, data=dat, ordered=names(dat), 
           test="mean.var.adjusted")

fitMeasures(fit, fit.measures = c("cfi.scaled","tli.scaled","rmsea.scaled",
                                  "rmsea.ci.lower.scaled","rmsea.ci.upper.scaled","rmsea.pvalue.scaled",
                                  "srmr") )

MI <- modificationIndices(fit)
MI[MI$sepc.lv>0.3,]


```

*Note: Check the model fit. If necessary improve model fit by allowing correlated errors cross-sectionally (e.g., `v1_1 ~~ v1_2, v2_1 ~~ v2_2`)*

### Reliability of TR

```{r}
pe <- parameterEstimates(fit, standardized=T, rsquare=T)
( rel.trat <- pe$est[pe$lhs=="trat" & pe$op=="r2"] )   # Reliability TR
```


## Estimate MICs

### ROC-based MIC

```{r}
xo1 <- rowSums(dat[,1:nitems])                 # sumscore T1
xo2 <- rowSums(dat[,(nitems+1):(2*nitems)])    # sumscore T2
dat$xoc <- xo2 - xo1                           # change score

( q <- mean(dat$trat) )            # q = proportion improved (perceived)
( p <- log(q/(1-q)) )              # p = logodds(pre)

( cor.trat.xoc <- cor(dat$trat, dat$xoc) )  # correlation anchor-change

## Do ROC analysis

rocobj <- roc(dat$trat, dat$xoc, quiet = TRUE)

( mic.roc <- coords(rocobj, x="best", input="threshold", ret="threshold", 
                    best.method="youden", transpose = TRUE) ) 

```


### predicted MIC

Apply a logistic regression for both the MIC(predicted) and MIC(adjusted)

```{r}

mylogit <- glm(trat ~ xoc, data = dat, family = "binomial")

C <- coef(mylogit)[1]                 # intercept coefficient C
B <- coef(mylogit)[2]                 # regression coefficient B

( mic.pred <- (p-C)/B )               # MIC(predicted) 
```


### Adjusted MIC

```{r}
rf <- (0.8/rel.trat - 0.5) * sd(dat$xoc) * cor.trat.xoc
( mic.adj <- mic.pred - rf * p )

```


## Bootstrap confidence intervals

### Bootstrap code and application

```{r}

nboot    <- 1000                # number of bootstrap samples
mic.roc  <- numeric(nboot)
mic.pred <- numeric(nboot)
mic.adj  <- numeric(nboot)
boot <- data.frame(mic.roc, mic.pred, mic.adj)

start.time <- Sys.time()

for(i in 1:nboot) {
  
  #print(i)
  
  dat <- org[sample(1:dim(dat)[1], dim(dat)[1], replace=TRUE),]
  
  ### Estimate TR reliability with lavaan
  
  # model is already defined in the previous section 
  
  fit <- cfa(model, data=dat, ordered=names(dat), 
             test="mean.var.adjusted")
  
  pe <- parameterEstimates(fit, standardized=T, rsquare=T)
  ( rel.trat <- pe$est[pe$lhs=="trat" & pe$op=="r2"] )   # Reliability TR
  
  ### Estimate MICs
  
  xo1 <- rowSums(dat[,1:nitems])                 # sumscore T1
  xo2 <- rowSums(dat[,(nitems+1):(2*nitems)])    # sumscore T2
  dat$xoc <- xo2 - xo1
  
  ( q <- mean(dat$trat) )            # q = proportion improved (perceived)
  ( p <- log(q/(1-q)) )              # p = logodds(pre)
  
  ( cor.trat.xoc <- cor(dat$trat, dat$xoc) )  # correlation anchor-change
  
  ## Do ROC analysis
  
  rocobj <- roc(dat$trat, dat$xoc, quiet = TRUE)
  
  mic.roc <- coords(rocobj, x="best", input="threshold", ret="threshold", 
                    best.method="youden", transpose = TRUE) 
  
  ( boot$mic.roc[i] <- mic.roc[sample(length(mic.roc),1)] )
  
  
  ## apply logistic regression and calculate parameters and MIC(pred)
  
  mylogit <- glm(trat ~ xoc, data = dat, family = "binomial")
  
  C <- coef(mylogit)[1]                 # intercept coefficient C
  B <- coef(mylogit)[2]                 # regression coefficient B
  
  ( boot$mic.pred[i] <- (p-C)/B )               # MIC(predicted) 
  
  ## Adjusted MIC
  rf <- (0.8/rel.trat - 0.5) * sd(dat$xoc) * cor.trat.xoc
  ( boot$mic.adj[i] <- boot$mic.pred[i] - rf * p )
  
}


end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```


### Bootstrap results

```{r}
round(mean(boot$mic.roc),1)
round(mean(boot$mic.pred),1)
round(mean(boot$mic.adj),1)

round(quantile(boot$mic.roc, c(0.025, 0.975)),1)
round(quantile(boot$mic.pred, c(0.025, 0.975)),1)
round(quantile(boot$mic.adj, c(0.025, 0.975)),1)

```
