---
title: "Chapter 6: MIC package demo"
output:
  pdf_document: default
  html_document: default
  
---
\setcounter{page}{19}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# R-package to estimate the adjusted MIC, predicted MIC and ROC-based MIC  

The R-package is public and can be installed from GitHub: `remotes::install_github(repo = "iriseekhout/MIC")`. 
```{r}
# to install MIC package from Github: 
#remotes::install_github(repo = "iriseekhout/MIC")

# Acquire packages
library(mirt)     # for simulating data
library(MIC)      # for calculating the MIC
``` 

## Simulate dataset using 'mirt'

This code simulates an example data set with 10 items that are measured at two time points. This dataset is used as an illustration.

### Gernerate a set of IRT item parameters

```{r}

set.seed(12345)

b2  <- c(-0.8, -0.8, -0.4, -0.4, 0, 0, 0.4, 0.4, 0.8, 0.8)
bc <- b2/4
b1 <- b2 - 1 + sample(bc)
b3 <- b2 + 1 + sample(bc)
a1 <- sample( 1+b2/2 )

( cf.simb <- data.frame(a1,b1,b2,b3) )
``` 


### Transform b-parameters to d-parameters (mirt works with d-parameters)

difficulty (b) = easiness (d) / -a

```{r}
cf.sim <- cf.simb

colnames(cf.sim) <- c("a1","d1","d2","d3")

cf.sim$d1 <- -cf.simb$b1*cf.sim$a1
cf.sim$d2 <- -cf.simb$b2*cf.sim$a1
cf.sim$d3 <- -cf.simb$b3*cf.sim$a1
round(cf.sim, 3)
round(colMeans(cf.sim), 3)
```



### Create theta T1 and dataset 1 (baseline)

```{r}
a1 <- as.matrix(cf.sim[ , 1])
d1 <- as.matrix(cf.sim[ , -1])

N <- 1000         # sample size
prop.imp <- 0.3   # proportion improved
mn.imic  <- 0.5   # mean individual MICs in terms of theta change

tet1s <- rnorm(N, 0, 1)   # simulate theta T1
tet1s <- as.matrix(tet1s)

set.seed( sample(10000:20000,1) )

dat1 <- simdata(a1, d1, N, itemtype="graded", Theta=tet1s)
dat1 <- as.data.frame(dat1)

xo1 <- rowSums(dat1)   # This is the baseline test (sum) score
```

### Create theta change

```{r}
rt1ch <- -0.5            # correlation between theta T1 and theta change
tetchs <- rt1ch*tet1s + sqrt(1-rt1ch^2)*rnorm(N, mean(tet1s), sd(tet1s))
cor(tet1s,tetchs)

tetchs <- tetchs - mean(tetchs)         # makes mean = 0
tetchs <- (tetchs/sd(tetchs))*1         # makes SD = 1

( qtl <- quantile(tetchs, prob=(1-prop.imp)) )
( mean.tetchs <- mn.imic - qtl )        # Estimate mean theta change to get
# the desired proportion improved
tetchs <- tetchs + mean.tetchs          # transform mean of tetchs
```

### Create theta T2 and dataset 2  (follow-up)

```{r}
tet2s <- as.matrix(tet1s + tetchs)     # theta T2

cor(tet1s, tet2s)   

set.seed( sample(30000:40000,1) )

dat2 <- simdata(a1, d1, N, itemtype="graded", Theta=tet2s)
dat2 <- as.data.frame(dat2)
```

### Insert error in transition ratings

```{r}

rel.trat <- 0.5   # reliability of the perceived change/transition rating

( sd.ch.error <- sqrt(((1-rel.trat)/rel.trat)*sd(tetchs)^2) )
tetch.percv <- tetchs + rnorm(N, 0, sd.ch.error)  # perceived change
```

### Create iMIC distribution
```{r}
imic <- rnorm(N, mn.imic, 0.1*mn.imic)
```

### create dichotomous transition ratings

```{r}
trat <- numeric(N)                         
trat[tetch.percv > imic] <- 1   
mean(trat)                           # proportion improved based on perceived change
```

### Create dataset

```{r}
org <- data.frame(dat1, dat2, trat)
head(org)
nitems = 10     # Provide the number of items in the scale

## Simplify/standardize the item names
names(org)[1:nitems] <- paste0('v1', '_', 1:nitems)
names(org)[(nitems+1):(2*nitems)] <- paste0('v2', '_', 1:nitems)
names(org)[2*nitems+1] <- "trat"
names(org)

## copy original data into a workfile (dat)
dat <- org
head(dat)

```




## Estimate TR reliability with lavaan

### Specify the factor model

A factor for each time point, with correlated errors over time.

```{r}
model <- ' 
   # Factors
   F1 =~ v1_1+v1_2+v1_3+v1_4+v1_5+v1_6+v1_7+v1_8+
         v1_9+v1_10+trat
   F2 =~ v2_1+v2_2+v2_3+v2_4+v2_5+v2_6+v2_7+v2_8+
         v2_9+v2_10+trat
   
   # Correlated errors over time
   v1_1 ~~ v2_1
   v1_2 ~~ v2_2
   v1_3 ~~ v2_3
   v1_4 ~~ v2_4
   v1_5 ~~ v2_5
   v1_6 ~~ v2_6
   v1_7 ~~ v2_7
   v1_8 ~~ v2_8
   v1_9 ~~ v2_9
   v1_10 ~~ v2_10
   '
```

### Use the `tr_reliability` function

```{r}
tr_reliability(example, model)
```

*Note: Check the model fit. If necessary improve model fit by allowing correlated errors cross-sectionally (e.g., `v1_1 ~~ v1_2, v2_1 ~~ v2_2`)*

#### Reliability of TR

```{r}
reliability <- tr_reliability(example, model, modification = FALSE)
```


## Estimate MICs

First prepare the sumscores in the data, to calculate the minimal important change for.

```{r}
nitems <- 10
example$x <- rowSums(example[,1:nitems])                 # sumscore T1
example$y <- rowSums(example[,(nitems+1):(2*nitems)])    # sumscore T2

```


### ROC-based MIC

```{r}
mic_roc(example, x = "x", y = "y", tr = "trat")
```


### predicted MIC

```{r}
mic_pred(example,x = "x", y = "y", tr = "trat")

```


### Adjusted MIC

```{r}
mic_adjust(example,x = "x", y = "y", tr = "trat", reliability = reliability)
```


## Bootstrap confidence intervals

### ROC-based MIC

```{r}

start <- Sys.time()
bootCImic(example, mic = "roc", x = "x",  y = "y", tr = "trat")
Sys.time() - start

```

### predicted MIC

```{r}

start <- Sys.time()
bootCImic(example, mic = "predict", x = "x",  y = "y", tr = "trat")
Sys.time() - start

```

### adjusted MIC

```{r}

start <- Sys.time()
bootCImic(example, mic = "adjust", x = "x",  y = "y", tr = "trat", model = model)
Sys.time() - start

```

